---
title: "Appendix"
author: "Alex Zwanenburg, Steffen Löck"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: lualatex
    keep_tex: true
papersize: a4
bibliography: refs.bib
header-includes:
- \usepackage{amsmath}
- \usepackage{amsthm}
- \usepackage{booktabs}
- \newtheorem{definition}{Definition}
- \DeclareMathOperator*{\argmax}{argmax}
- \DeclareMathOperator*{\argmin}{argmin}
- \DeclareMathOperator{\sgn}{sgn}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  eval = TRUE,
  fig.align = "center",
  dev = "cairo_pdf"
)

# Allow for defining figure captions within a chunk.
knitr::opts_knit$set(
  eval.after = "fig.cap")

require(power.transform)
require(ggplot2)
require(patchwork)
require(data.table)
require(latex2exp)
require(coro)
require(paletteer)
require(robustHD)

manuscript_dir <- file.path("C:/Users/alexz/Documents/GitHub/power.transform/manuscript_rmarkdown")
source(file.path(manuscript_dir, "simulations.R"))
source(file.path(manuscript_dir, "plots.R"))

# Base font size.
base_size <- 9

# Set general theme.
plot_theme <- ggplot2::theme_light(base_size=base_size)
plot_theme$plot.title$size <- ggplot2::rel(1.0)
plot_theme$plot.title$hjust <- 0.5
plot_theme$strip.background <- ggplot2::element_blank()
plot_theme$strip.text$colour <- NULL
plot_theme$plot.margin <- grid::unit(c(2, 2, 2, 2), "points")
plot_theme$legend.key.size <- grid::unit(0.3, 'lines')
plot_theme$plot.tag <- ggplot2::element_text(
  size = ggplot2::rel(1.0),
  hjust = 0.0,
  vjust = 0.0,
  inherit.blank = FALSE
)
plot_theme$plot.tag.position <- c(0.2, 1)

parse_mean_sd_latex <- function(mu, sigma, digits=2L){
  return(paste0(
    "$", round(mu, digits=digits), " \\pm ", round(sigma, digits=digits), "$"
  ))
}

```

# Appendix A: Log-likelihood functions for location and scale invariant power transformation

Location and scale-invariant Box-Cox and Yeo-Johnson transformations are parametrised using location $x_0$ and scale $s$ parameters, in addition to transformation parameter $\lambda$.
This leads to the following transformations. The location and scale-invariant Box-Cox transformation is:

```{=latex}
\begin{equation}
\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) = 
\begin{cases}
\left( \left(\frac{x_i - x_0}{s} \right)^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
\log\left[\frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0
\end{cases}
\end{equation}
```
where $x_i - x_0 > 0$.
The location and scale-invariant Yeo-Johnson transformation is:

```{=latex}
\begin{equation}
\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) = 
\begin{cases}
\left( \left( 1 + \frac{x_i - x_0}{s}\right)^\lambda - 1\right) / \lambda & \text{if } \lambda \neq 0 \text{ and } x_i - x_0 \geq 0\\
\log\left[1 + \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 0 \text{ and } x_i - x_0 \geq 0\\
-\left( \left( 1 - \frac{x_i - x_0}{s}\right)^{2 - \lambda} - 1 \right) / \left(2 - \lambda \right) & \text{if } \lambda \neq 2 \text{ and } x_i - x_0 < 0\\
-\log\left[1 - \frac{x_i - x_0}{s}\right] & \text{if } \lambda = 2 \text{ and } x_i - x_0 < 0
\end{cases}
\end{equation}
```

The parameters of these power transformations can be optimised based by maximising the log-likelihood function, under the assumption that the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ follows a normal distribution.
The log-likelihood functions for conventional Box-Cox and Yeo-Johnson transformations are well-known.
However, the introduction of scaling parameter $s$ prevents their direct use.
Here, we first derive the general form of the log-likelihood functions, and then derive their power-transformation specific definitions.

Let $f(x_1, \ldots, x_n)$ be the probability density function of feature $\mathbf{X} = \{ x_1, \ldots, x_n\}$, 
and $f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n))$ be the probability density function of 
the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$, that is assumed to follow a normal distribution.

The two probability density functions are related as follows:

```{=latex}
\begin{equation}
f^{\lambda, x_0, s}(x_1, \ldots, x_n) = f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \left|\mathbf{J}\right|
\end{equation}
```
Where, $\left|\mathbf{J}\right|$ is the determinant of Jacobian $\mathbf{J}$.
The Jacobian takes the following form, with off-diagonal elements $0$:
```{=latex}
\begin{equation}
\mathbf{J} =
\begin{bmatrix}
    \frac{\partial}{\partial x_1} \phi^{\lambda, x_0, s}(x_1) & 0 & \dots & 0 \\
    0 & \frac{\partial}{\partial x_2} \phi^{\lambda, x_0, s}(x_2) & \dots & 0 \\
    \vdots & \vdots  & \ddots &  \vdots \\
    0  & 0 & 0 & \frac{\partial}{\partial x_n} \phi^{\lambda, x_0, s}(x_n)
\end{bmatrix}
\end{equation}
```

Thus, $\left| \mathbf{J} \right| = \prod_{i=1}^n \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i)$.

Since in our situation $\{x_1, \ldots, x_n\}$ in $f^{\lambda, x_0, s}(x_1, \ldots, x_n)$ are considered fixed (i.e., known), $f^{\lambda, x_0, s}(x_1, \ldots, x_n)$ may be considered a likelihood function.
The log-likelihood function $\mathcal{l}^{\lambda, x_0, s}$ is then:
```{=latex}
\begin{equation}
\begin{split}
\mathcal{l}^{\lambda, x_0, s} & = \log f^{\lambda, x_0, s}(x_1, \ldots, x_n) \\
 & = \log \left[ f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \right] + \log \left|\mathbf{J}\right| \\
 & = \log \left[ f^{\lambda, x_0, s} (\phi^{\lambda, x_0, s}(x_1), \ldots, \phi^{\lambda, x_0, s}(x_n)) \right] + \log \prod_{i=1}^n \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i) \\
 & = -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 + \sum_{i=1}^n \log \left[ \frac{\partial}{\partial x_i} \phi^{\lambda, x_0, s}(x_i)\right]
\end{split}
\end{equation}
```
With $\mu$ the average of $\phi^{\lambda, x_0, s}(\mathbf{X})$ and $\sigma^2$ its variance.
The first two terms derive directly from the log-likelihood function of a normal distribution, and are not specific to the type of power transformation used.
However, the final term differs between Box-Cox and Yeo-Johnson transformations.

## Location- and scale-invariant Box-Cox transformation

For the location- and scale-invariant Box-Cox transformation the partial derivative is:
```{=latex}
\begin{equation}
\begin{split}
\frac{\partial}{\partial x_i} \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) & = \frac{1}{s} \left(\frac{x_i - x_0}{s} \right)^{\lambda-1} \\
 & = \frac{1} {s^\lambda} \left(x_i - x_0 \right)^{\lambda - 1}
\end{split}
\end{equation}
```

Thus the final term in $\mathcal{l}_{\text{BC}}^{\lambda, x_0, s}$ is:
```{=latex}
\begin{equation}
\begin{split}
\sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi_{\text{BC}}^{\lambda, x_0, s}(x_i) & = \sum_{i=1}^n \log \left[ s^{-\lambda} (x_i - x_0)^{\lambda - 1} \right] \\
& = \sum_{i=1}^n \log \left[s^{-\lambda} \right] + \log \left[ (x_i - x_0)^{\lambda - 1} \right]\\
& = -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
\end{split}
\end{equation}
```

This leads to the following log-likelihood:
```{=latex}
\begin{equation}
\begin{split}
\mathcal{l}_{\text{BC}}^{\lambda, x_0, s} = & -\frac{n}{2} \log \left[2 \pi \sigma^2 \right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
& -n \lambda \log s + \left( \lambda - 1 \right) \sum_{i=1}^n \log \left[ x_i - x_0 \right]
\end{split}
\end{equation}
```

Similarly to @Raymaekers2024-zf, sample weights $w_i$ are introduced to facilitate robust power transformations.
The weighted log-likelihood of the location- and scale-invariant Box-Cox transformation is:
```{=latex}
\begin{equation}
\begin{split}
\mathcal{l}_{\text{rBC}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
& - \lambda \left( \sum_{i=1}^n w_i \right) \log s + \left( \lambda - 1 \right) \sum_{i=1}^n w_i \log \left[ x_i - x_0 \right]
\end{split}
\end{equation}
```
where $\mu_w$ and $\sigma^2_w$ are the weighted mean and weighted variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0, s} (\mathbf{X})$, respectively:

```{=latex}
\begin{equation}
\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{BC}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{BC}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
\end{equation}
```

## Location- and scale-invariant Yeo-Johnson transformation

For the location- and scale-invariant Yeo-Johnson transformation, the partial derivative is:
```{=latex}
\begin{equation}
\frac{\partial}{\partial x_i} \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) =
\begin{cases}
\frac{1}{s} \left(1 + \frac{x_i - x_0}{s}\right)^{\lambda - 1} & \text{if } x_i - x_0 \geq 0\\
\frac{1}{s} \left(1 - \frac{x_i - x_0}{s}\right)^{1 - \lambda} & \text{if } x_i - x_0 < 0
\end{cases}
\end{equation}
```

Thus the final term in $\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s}$ is:
```{=latex}
\begin{equation}
\begin{split}
\sum_{i=1}^n \log \frac{\partial}{\partial x_i} \phi_{\text{YJ}}^{\lambda, x_0, s}(x_i) & = - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
\end{split}
\end{equation}
```

This leads to the following log-likelihood:
```{=latex}
\begin{equation}
\begin{split}
\mathcal{l}_{\text{YJ}}^{\lambda, x_0, s} = & -\frac{n}{2} \log\left[2 \pi \sigma^2\right] -\frac{1}{2 \sigma^2} \sum_{i=1}^n \left( \phi^{\lambda, x_0, s}(x_i) - \mu \right)^2 \\
& - n \log s + (\lambda - 1) \sum_{i=1}^n \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
\end{split}
\end{equation}
```

The weighted log-likelihood for location- and scale-invariant Yeo-Johnson transformation is:
```{=latex}
\begin{equation}
\begin{split}
\mathcal{l}_{\text{rYJ}}^{\lambda, x_0, s} = & -\frac{1}{2} \left(\sum_{i=1}^n w_i \right) \log \left[ 2 \pi \sigma_w^2 \right] -\frac{1}{2 \sigma_w^2} \sum_{i=1}^n w_i \left( \phi^{\lambda, x_0, s}(x_i) - \mu_w \right)^2 \\
& - \left( \sum_{i=1}^n w_i \right) \log s + (\lambda - 1) \sum_{i=1}^n w_i \sgn(x_i - x_0) \log \left[1 + \frac{|x_i - x_0|}{s} \right]
\end{split}
\end{equation}
```
where $\mu_w$ and $\sigma^2_w$ are the weighted mean and weighted variance of the Yeo-Johnson transformed feature $\phi_{\text{YJ}}^{\lambda, x_0, s} (\mathbf{X})$:

```{=latex}
\begin{equation}
\sigma_w^2 = \frac{\sum_{i=1}^n w_i \left(\phi_{\text{YJ}}^{\lambda, x_0, s} (x_i) - \mu_w \right)^2}{\sum_{i=1}^n w_i} \quad \text{with } \mu_w = \frac{\sum_{i=1}^n \phi_{\text{YJ}}^{\lambda, x_0, s} (x_i)} {\sum_{i=1}^n w_i}
\end{equation}
```


# Appendix B: Optimisation of transformation parameters

Maximum likelihood estimation (MLE) is commonly used to optimise parameters for power transformation.
Generally, optimisation requires minimisation or maximisation of a criterion.
In MLE, the maximised criterion is the log-likelihood function of the normal distribution.
Here, we investigate power transformation using optimisation criteria that are closely related to test statistics for normality tests.

Let $\mathbf{X}$ be a feature with ordered feature values, and $\mathbf{Y}^\lambda =\phi^{\lambda} \left(\mathbf{X} \right)$ and $\mathbf{Y}^{\lambda, x_0, s} =\phi^{\lambda, x_0, s} \left(\mathbf{X} \right)$ its transformed values using conventional and shift and scale invariant power transformations, respectively. Since power transformations are monotonic, $\mathbf{Y}$ will likewise be ordered.

Below we will focus on criteria based on the empirical density function and those based on skewness and kurtosis of the transformed featured.
Other potential criteria, such as the Shapiro-Wilk test statistic [@Shapiro1965-zd] are not investigated here.
In the case of the Shapiro-Wilk test statistic this is because of lack of scalability to features with many ($> 5000$) instances, and because adapting the test statistic to include weights is not straightforward.

## Empirical density function-based criteria

The first class of criteria is based on the empirical distribution function (EDF).
Transformation parameters are then fit through minimisation of the distance between the empirical distribution function $F_{\epsilon}$ and the cumulative density function (CDF) of the normal distribution $F_{\mathcal{N}}$.
Let $F_{\epsilon}\left(x_i \right) = \frac{i - 1/3}{n + 1/3}$ be the empirical probability of instance $i$.
The normal distribution is parametrised by location parameter $\mu$ and scale parameter $\sigma$, both of which have to be estimated from the data. 
For non-robust power transformations, $\mu$ and $\sigma$ are sample mean and sample standard deviation, respectively.
For robust power transformations, we estimate $\mu$ and $\sigma$ as Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0, s} (\mathbf{X})$ [@Huber1981-su].

### Anderson-Darling criterion

The Anderson-Darling criterion is based on the empirical distribution function of $\mathbf{X}$.
We define this criterion as follows:

```{=latex}
\begin{equation}
U_{\text{AD}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \frac{\left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\phi^{\lambda, x_0, s} \left(x_i \right); \mu, \sigma \right) \right)^2} {F_{\mathcal{N}} \left(\phi^{\lambda, x_0, s} \left(x_i \right); \mu, \sigma \right) \left(1 - F_{\mathcal{N}} \left(\phi^{\lambda, x_0, s} \left(x_i \right); \mu, \sigma \right) \right) }
\end{equation}
```

Here $w_i$ are weights, and $\mu$ and $\sigma$ are location and scale parameters. 
For non-robust power transformations, all $w_i = 1$.
Note that this criterion is not the same as the Anderson-Darling test statistic [@Anderson1952-gz], which involves solving (or approximating) an integral function, contains an extra scalar multiplication term, and does not include weights.
The Anderson-Darling criterion seeks to minimise the squared Euclidean distance between the EDF and the normal CDF, with differences at the upper and lower end of the normal CDF receiving more weight than those at the the centre of the CDF.

### Cramér-von Mises criterion

The Cramér-von Mises criterion is also based on the empirical distribution function of $\mathbf{X}$.
We define the Cramér-von Mises criterion as follows:
```{=latex}
\begin{equation}
U_{\text{CvM}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\phi^{\lambda, x_0, s} \left(x_i \right); \mu, \sigma \right) \right)^2
\end{equation}
```

Here $w_i$ are weights, and $\mu$ and $\sigma$ are location and scale parameters.
For non-robust power transformations, all $w_i = 1$.
The criterion is similar to the Cramér-von Mises test statistic [@Cramer1928-rc, @Von_Mises1928-ef], aside from a additive scalar value and the introduction of weights.
This criterion, like the Anderson-Darling criterion, seeks to minimise the squared Euclidean distance between the EDF and the normal CDF.
Unlike the Anderson-Darling criterion, this criterion weights all instances equally.

For conventional power transformations with a fixed shift parameter, the transformation $\phi^{\lambda, x_0, s} (\mathbf{X})$ may be substituted by $\phi^{\lambda} (\mathbf{X})$ in the definition of the Cramér-von Mises criterion.

## Skewness-kurtosis-based criteria

The second class of criteria seeks to reduce skewness and (excess) kurtosis of the transformed feature $\mathbf{Y}$.
We will first define the location $\mu$ and scale $\sigma$ of the the transformed as these are required for computing skewness and kurtosis.
Here, $\mu$ is defined as:

```{=latex}
\begin{equation}
\mu = \frac{\sum_{i=1}^n \phi^{\lambda, x_0, s} \left(x_i \right)} {\sum_{i=1}^n w_i}
\end{equation}
```

The location, or mean, is weighted using weights $w_i$.
For non-robust transformations, $w_i = 1$.
Then, $\sigma^2$ is defined as:
```{=latex}
\begin{equation}
\sigma^2 = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0, s} \left( x_i \right) - \mu \right)^2}{\sum_{i=1}^n w_i}
\end{equation}
```

Skewness is defined as:
```{=latex}
\begin{equation}
s = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0, s} \left( x_i \right) - \mu \right)^3}{\sigma^3 \sum_{i=1}^n w_i}
\end{equation}
```

Kurtosis is defined as:
```{=latex}
\begin{equation}
k = \frac{\sum_{i=1}^n w_i \left(\phi^{\lambda, x_0, s} \left( x_i \right) - \mu \right)^4}{\sigma^4 \sum_{i=1}^n w_i}
\end{equation}
```

### D'Agostino criterion

The D'Agostino criterion defined here follows the D'Agostino $K^2$ test statistic [@DAgostino1990-kp].
This test statistic is composed of two separate test statistics, one of which is related to skewness, and the other to kurtosis.
Both test statistics are computed in several steps.
Let us first define $\nu=\sum_{i=1}^n w_i$.
Thus for non-robust power transformations, $\nu = n$.

For the skewness test statistic we first compute [@DAgostino1990-kp]:

```{=latex}
\begin{equation}
\beta_1 = s \sqrt{ \frac{\left(\nu + 1\right) \left(\nu + 3\right)} {6 \left(\nu - 2\right)} }
\end{equation}

\begin{equation}
\beta_2 = 3 \frac{\left(\nu^2 + 27\nu - 70\right) \left(\nu + 1\right) \left(\nu + 3\right)} {\left(\nu - 2\right) \left(\nu + 5\right) \left(\nu + 7\right) \left(\nu + 9\right)}
\end{equation}

\begin{equation}
\alpha = \sqrt{\frac{2} {\sqrt{2 \beta_2 - 2} - 2}}
\end{equation}

\begin{equation}
\delta = \frac{1}{\sqrt{\log \left[\sqrt{-1 + \sqrt{2 * \beta_2 - 2}} \right]}}
\end{equation}
```

The skewness test statistic is then:

```{=latex}
\begin{equation}
Z_s = \delta \log\left[\frac{\beta_1}{\alpha} + \sqrt{\frac{\beta_1^2}{\alpha^2} + 1} \right]
\end{equation}
```

For the kurtosis test statistic we first compute [@DAgostino1990-kp; @Anscombe1983-nz]:

```{=latex}
\begin{equation}
\beta_1 = 3 \frac{\nu - 1}{\nu + 1}
\end{equation}

\begin{equation}
\beta_2 = 24 \nu \frac{\left(\nu - 2\right)\left(\nu - 3\right)}{\left(\nu + 1\right)^2 \left(\nu + 3\right) \left(\nu + 5\right)}
\end{equation}

\begin{equation}
\beta_3 = 6 \frac{\nu^2 - 5 \nu + 2}{\left(\nu + 7\right) \left(\nu + 9\right)} \sqrt{6 \frac{\left(\nu + 3\right) \left(\nu + 5\right)}{\nu \left(\nu - 2\right) \left(\nu - 3 \right)}}
\end{equation}

\begin{equation}
\alpha_1 = 6 + \frac{8}{\beta_3} \left[\frac{2}{\beta_3} + \sqrt{1 + \frac{4}{\beta_3^2}} \right]
\end{equation}

\begin{equation}
\alpha_2 = \frac{k - \beta_1}{\sqrt{\beta_2}}
\end{equation}
```

The kurtosis test statistic is then:

```{=latex}
\begin{equation}
Z_k = \sqrt{\frac{9 \alpha_1}{2}} \left[ 1 - \frac{2}{9 \alpha_1} - \left(\frac{1 - 2 / \alpha_1}{1 + \alpha_2 \sqrt{2 / \left(\alpha_1 - 4 \right)}} \right)^{1 / 3}  \right]
\end{equation}
```

The D'Agostino $K^2$ test statistic and our criterion are the same, and are defined as:
```{=latex}
\begin{equation}
U_{\text{DA}} \left(\mathbf{X}, \lambda, x_0 \right) = Z_s^2 + Z_k^2
\end{equation}
```

The main difference between the test statistic as originally formulated, and the criterion proposed here is the presence of weights for robust power transformation.

### Jarque-Bera criterion

The second criterion based on skewness and kurtosis is the Jarque-Bera criterion.
It is relatively simple to compute compared to the D'Agostino criterion:
```{=latex}
\begin{equation}
U_{\text{JB}} \left(\mathbf{X}, \lambda, x_0 \right) = s^2 + \left(k - 3\right)^2 / 4
\end{equation}
```

The main difference between the above criterion and the Jarque-Bera test statistic [@Jarque1980-hw] is that a scalar multiplication is absent.

## Optimisation using non-MLE criteria

Each of the above criteria can be used for optimisation, i.e.:

```{=latex}
\begin{equation}
\left\{ \hat{\lambda}, \hat{x}_0, \hat{s}_0 \right\} = \argmin_{\lambda, x_0, s} U\left(\mathbf{X}, \lambda, x_0, s \right)
\end{equation}
```

For conventional power transformations with fixed location and scale parameters, the transformation $\phi^{\lambda, x_0, s} (\mathbf{X})$ may be substituted by $\phi^{\lambda} (\mathbf{X})$, or equivalently, $x_0$ and $s$ may be fixed:
```{=latex}
\begin{equation}
\left\{ \hat{\lambda}\right\} = \argmin_{\lambda} U\left(\mathbf{X}, \lambda; x_0, s \right)
\end{equation}
```


# Appendix C: Simulations with other optimisation criteria

Invariance of location- and scale-invariant power transformations was assessed using the optimisation criteria in [Appendix B](Appendix B: Optimisation of transformation parameters).
This follows the simulation in the main manuscript, where MLE was used for optimization.
In short, we first randomly drew $10000$ values from a normal distribution: $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{N}\left(0, 1\right)$, or equivalently $\mathbf{X}_{\text{normal}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.5, 2\right)$.
The second distribution was a right-skewed normal distribution $\mathbf{X}_{\text{right}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.2, 2\right)$.
The third distribution was a left-skewed normal distribution $\mathbf{X}_{\text{left}} = \left\{x_1, x_2, \ldots, x_{10000} \right\} \sim \mathcal{AGN}\left(0, 1/\sqrt{2}, 0.8, 2\right)$.
We then computed transformation parameter $\lambda$ using the original definitions (equations \ref{eqn:box-cox-original} and \ref{eqn:yeo-johnson-original}) and the location- and scale-invariant definitions (equations \ref{eqn:box-cox-invariant} and \ref{eqn:yeo-johnson-invariant}) for each distribution using different optimisation criteria.
To assess location invariance, a positive value $d_{\text{shift}}$ was added to each distribution with $d_{\text{shift}} \in [1, 10^6]$.
Similarly, to assess scale invariance, each distribution was multiplied by a positive value $d_{\text{scale}}$, where $d_{\text{scale}} \in [1, 10^6]$.

The results are shown in Figure \ref{fig:shifted-distributions-appendix}.

```{r shifted-distributions-appendix, warning=FALSE, message=FALSE, fig.cap=cap}
cap <- paste0(
  "Invariant power transformation produces transformation parameters that are invariant to location and scale. ",
  "Samples were drawn from normal, right-skewed and left-skewed distributions, respectively, which then underwent a shift $d_{\\text{shift}}$ or multiplication by $d_{\\text{scale}}$. ",
  "Estimates of the transformation parameter $\\lambda$ for the conventional power transformations show strong dependency on the overall location and scale of the distribution and the optimisation criterion, ",
  "whereas estimates obtained for the location- and scale-invariant power transformations are constant. ",
  "For location- and scale-invariant power transformations, the Anderson-Darling criterion leads to unstable estimates of $\\lambda$ for skewed distributions, possibly due to large weights being assigned to samples at the upper and lower ends of the distribution." 
)

.plot_shifted_distributions_appendix(manuscript_dir = manuscript_dir, plot_theme = plot_theme)
```

# Appendix D: Experimental results using location- and scale-invariant Box-Cox transformation

The effect of using location- and scale-invariant transformations was investigated using real-world datasets.

## Invariance

Results for Box-Cox transformations of features without outliers are shown in Figure \ref{fig:experimental-results-invariance-appendix}.

```{r experimental-results-invariance-appendix, warning=FALSE, message=FALSE, fig.cap=cap}
cap <- paste0(
  "Quantile-quantile plots for several datasets: age of patients with lung cancer (top row); ",
  "penguin body mass (middle row); ",
  "and latitude coordinates of houses sold in Ames, Iowa (bottom row). ",
  "Multiple quantile-quantile plots are shown: for the original feature (left column); ",
  "the feature transformed using the conventional Box-Cox transformation ",
  "and Raymaekers and Rousseeuw's robust adaptation (middle column); ",
  "and the feature transformed using the non-robust and robust location- and-scale ",
  "invariant Box-Cox transformations (right column)."
)

data <- .plot_experimental_invariance(plot_theme = plot_theme, method = "box_cox")
data$plot
```

### Age of patients with lung cancer

Applying conventional and invariant Box-Cox transformations to age of patients with lung cancer [@Loprinzi1994-cd] yielded the following results:
no transformation (sum of residuals with normal distribution $\sum r_i = 16.5$, $p = 0.71$);
conventional transformation ($\lambda = 1.9$, $\sum r_i = 11.5$, $\mu_{BC} = 1.6 \cdot 10^3$, $\sigma_{BC} = 0.4 \cdot 10^3$, $p = 0.94$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.9$, $\sum r_i = 11.5$, $\mu_{BC} = 1.6 \cdot 10^3$, $\sigma_{BC} = 0.4 \cdot 10^3$, $p = 0.94$);
location- and scale-invariant transformation ($\lambda = 1.7$, $\sum r_i = 11.6$, $\mu_{BC} = 1.9$, $\sigma_{BC} = 0.8$, $p = 0.94$);
and robust location- and scale-invariant transformation ($\lambda = 1.5$, $\sum r_i = 11.6$, $\mu_{BC} = 3.6$, $\sigma_{BC} = 1.2$, $p = 0.93$).

Compared to location- and scale-invariant Yeo-Johnson transformations, the Box-Cox transformations do not reduce residuals compared to conventional variants.


### Penguin body mass

Applying conventional and invariant Box-Cox transformations to the body mass of penguins [@Gorman2014-eo] yielded the following results:
no transformation (residual sum $\sum r_i = 48.0$, $p < 0.001$);
conventional transformation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{BC} = 2.1$, $\sigma_{BC} = 4 \cdot 10^{-3}$, $p = 0.06$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = -0.5$, $\sum r_i = 32.2$, $\mu_{BC} = 2.1$, $\sigma_{BC} = 4 \cdot 10^{-3}$, $p = 0.06$);
location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 27.3$, $\mu_{BC} = 0.3$, $\sigma_{BC} = 0.6$, $p = 0.10$);
and robust location- and scale-invariant transformation ($\lambda = 0.2$, $\sum r_i = 25.2$, $\mu_{BC} = 0.4$, $\sigma_{BC} = 0.6$, $p = 0.44$).

Just as for location- and scale-invariant Yeo-Johnson transformations, Box-Cox transformations produced a lower overall residual sum compared to their conventional counterparts.
Similarly, conventional transformations led to low standard deviation $\sigma_{YJ}$ of the body mass feature after transformation.


### Latitude in the Ames housing dataset

Applying conventional and invariant Box-Cox transformations to the latitude of houses in the Ames housing dataset [@De-Cock2011-jf] yielded the following results:
no transformation (residual sum $\sum r_i = 328$, $p < 0.001$);
conventional transformation ($\lambda = 62.1$, $\sum r_i = 319$, $\mu_{BC} = 1.1 \cdot 10^{99}$, $\sigma_{BC} = 0.0 \cdot 10^{99}$, $p < 0.001$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = 96.0$, $\sum r_i = 319$, $\mu_{BC} = 6.2 \cdot 10^{153}$, $\sigma_{BC} = 0.3 \cdot 10^{153}$, $p < 0.001$);
location- and scale-invariant transformation ($\lambda = 1.9$, $\sum r_i = 312$, $\mu_{BC} = 2.3$, $\sigma_{BC} = 0.9$, $p < 0.001$);
and robust location- and scale-invariant transformation ($\lambda = 1.2$, $\sum r_i = 316$, $\mu_{BC} = 5.5$, $\sigma_{BC} = 1.4$, $p < 0.001$).

Similar to conventional Yeo-Johnson transformations (non-robust and robust), conventional Box-Cox transformations had high values for the $\lambda$ parameter, which could lead to numerical issues.
Location- and scale-invariant Box-Cox transformations did not suffer from this issue.


## Robustness against outliers

Results for Box-Cox transformations of features with outliers are shown in Figure \ref{fig:experimental-results-outlier-robustness-appendix}.

```{r experimental-results-outlier-robustness-appendix, warning=FALSE, message=FALSE, fig.cap=cap}
cap <- paste0(
  "Quantile-quantile plots for two datasets with outliers: vehicle fuel consumption ",
  "(top row), where outliers are related to highly fuel-efficient vehicles; ",
  "and maximum arterial wall thickness in patients with ischemic stroke (bottom row). ",
  "Multiple quantile-quantile plots are shown: for the original feature (left column); ",
  "the feature transformed using the conventional Box-Cox transformation ",
  "and Raymaekers and Rousseeuw's robust adaptation (middle column); ",
  "and the feature transformed using the non-robust and robust location- and-scale ",
  "invariant Box-Cox transformations (right column). ",
  "Samples with observed quantiles below $-3.0$ or above $3.0$ are indicated by crosses."
)

data <- .plot_experimental_outlier_robustness(plot_theme = plot_theme, method = "box_cox")
data$plot
```

### Fuel efficiency in the Top Gear dataset

The Top Gear dataset contains data on 297 vehicles, with outliers related highly fuel-efficient vehicles [@Alfons2021-kc].
Applying conventional and invariant Box-Cox transformations to the fuel consumption feature yielded the following results:
no transformation (residual sum $\sum r_i = 54$, $p=0.58$);
conventional transformation ($\lambda = -0.1$, $\sum r_i = 55$, $\mu_{BC} = 3.0$, $\sigma_{BC} = 0.3$, $p < 0.001$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = 0.8$, $\sum r_i = 48$, $\mu_{BC} = 29$, $\sigma_{BC} = 15$, $p=0.32$);
location- and scale-invariant transformation ($\lambda = -0.7$, $\sum r_i = 44$, $\mu_{BC} = 0.6$, $\sigma_{BC} = 0.2$, $p < 0.001$);
and robust location- and scale-invariant transformation ($\lambda = 1.1$, $\sum r_i = 59$, $\mu_{BC} = 2.4$, $\sigma_{BC} = 1.8$, $p=0.67$).


### Maximum arterial wall thickness in an ischemic stroke dataset

The ischemic stroke dataset contains historic data from 126 patients with risk at ischemic stroke [@Kuhn2019-kt].
Applying conventional and invariant Box-Cox transformations to the maximum arterial wall thickness feature yielded the following results:
no transformation (residual sum $\sum r_i = 110$, $p=0.86$);
conventional transformation ($\lambda = -0.5$, $\sum r_i = 33$, $\mu_{BC} = 1.0$, $\sigma_{BC} = 0.2$, $p=0.01$);
Raymaekers and Rousseeuw's robust adaptation ($\lambda = 1.1$, $\sum r_i = 127$, $\mu_{BC} = 5.5$, $\sigma_{BC} = 12$, $p=0.90$);
location- and scale-invariant transformation ($\lambda = -1.0$, $\sum r_i = 28$, $\mu_{BC} = 0.7$, $\sigma_{BC} = 0.1$, $p=0.02$);
and robust location- and scale-invariant transformation ($\lambda = 0.5$, $\sum r_i = 56$, $\mu_{BC} = 2.2$, $\sigma_{BC} = 1.4$, $p=0.60$).

# Appendix E: Empirical central normality test

Critical values for the central normality test and its empirical variant are
found in Table \ref{tab:central-normality-critical-statistic} and Table 
\ref{tab:empirical-central-normality-critical-statistic}, respectively.

```{r include = FALSE, eval = FALSE}
# This writes the lookup table for critical test statistic.
load(file = "./R/sysdata.rda")

central_normality_lookup_table <- .get_test_statistic_lookup_table(
  manuscript_dir = manuscript_dir,
  k = 0.70,
  for_manuscript = FALSE
)

empirical_central_normality_lookup_table <- .get_test_statistic_lookup_table(
  manuscript_dir = manuscript_dir,
  k = 0.70,
  for_manuscript = FALSE,
  with_outliers = TRUE
)

usethis::use_data(
  central_normality_lookup_table,
  empirical_central_normality_lookup_table,
  two_sided_function_parameters, 
  internal = TRUE, 
  overwrite = TRUE)
```

```{r include = FALSE, eval = FALSE}
# This generates the table shown below.
.get_test_statistic_lookup_table(
  manuscript_dir = manuscript_dir,
  k = 0.70,
  for_manuscript = TRUE
)

```

```{=latex}
\begin{table}
\begin{center}
\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
\textbf{central normality} at $\kappa = 0.70$, as a function of significance 
level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
\label{tab:central-normality-critical-statistic}
\begin{tabular}{l | r r r r r r r r r}

\toprule
$n$ \textbackslash $\alpha$ & 0.9 & 0.8 & 0.5 & 0.2 & 0.1 & 0.05 & 0.025 & 0.01 & 0.001 \\

\midrule
5     & 10.49 & 13.15 & 17.46 & 22.17 & 25.28 & 27.90 & 29.93 & 31.88 & 34.58 \\
10    &  8.67 & 10.14 & 13.66 & 18.16 & 21.11 & 23.91 & 26.96 & 31.44 & 48.04 \\
20    &  6.77 &  7.71 &  9.97 & 12.96 & 14.93 & 16.70 & 18.24 & 20.33 & 25.24 \\
50    &  4.63 &  5.20 &  6.59 &  8.49 &  9.69 & 10.81 & 11.87 & 13.14 & 16.11 \\
100   &  3.32 &  3.73 &  4.71 &  6.06 &  6.92 &  7.71 &  8.45 &  9.39 & 11.31 \\
200   &  2.37 &  2.66 &  3.35 &  4.29 &  4.89 &  5.46 &  6.00 &  6.67 &  8.02 \\
500   &  1.51 &  1.69 &  2.12 &  2.73 &  3.12 &  3.48 &  3.84 &  4.24 &  5.30 \\
1000  &  1.07 &  1.20 &  1.50 &  1.92 &  2.19 &  2.44 &  2.68 &  3.01 &  3.68 \\
2000  &  0.76 &  0.85 &  1.06 &  1.37 &  1.56 &  1.74 &  1.90 &  2.11 &  2.55 \\
5000  &  0.48 &  0.54 &  0.67 &  0.86 &  0.99 &  1.10 &  1.20 &  1.34 &  1.64 \\
10000 &  0.34 &  0.38 &  0.48 &  0.61 &  0.69 &  0.77 &  0.84 &  0.93 &  1.15 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
```

```{r include = FALSE, eval = FALSE}
# This generates the table shown below.
.get_test_statistic_lookup_table(
  manuscript_dir = manuscript_dir,
  k = 0.70,
  for_manuscript = TRUE,
  with_outliers = TRUE
)

```

```{=latex}
\begin{table}
\begin{center}
\caption{Critical values of test statistic $\tau_{\alpha, n, \kappa = 0.70}$ for
\textbf{empirical central normality} at $\kappa = 0.70$, as a function of significance 
level $\alpha$ and number of instances $n$. The values shown must be divided by $100$.}
\label{tab:empirical-central-normality-critical-statistic}
\begin{tabular}{l | r r r r r r r r r}

\toprule
$n$ \textbackslash $\alpha$ & 0.9 & 0.8 & 0.5 & 0.2 & 0.1 & 0.05 & 0.025 & 0.01 & 0.001 \\

\midrule
5     & 12.04 & 14.02 & 17.37 & 21.26 & 24.02 & 26.45 & 28.46 & 30.65 & 33.90 \\
10    &  8.75 & 10.23 & 13.82 & 18.53 & 21.58 & 24.31 & 27.01 & 30.82 & 41.82 \\
20    &  6.80 &  7.74 & 10.03 & 13.17 & 15.12 & 16.95 & 18.64 & 20.79 & 25.07 \\
50    &  4.67 &  5.26 &  6.69 &  8.70 &  9.99 & 11.18 & 12.32 & 13.65 & 16.73 \\
100   &  3.38 &  3.79 &  4.83 &  6.21 &  7.13 &  7.97 &  8.72 &  9.72 & 11.78 \\ 
200   &  2.47 &  2.77 &  3.51 &  4.54 &  5.19 &  5.77 &  6.31 &  6.92 &  8.47 \\
500   &  1.65 &  1.86 &  2.38 &  3.06 &  3.47 &  3.85 &  4.19 &  4.62 &  5.58 \\
1000  &  1.27 &  1.44 &  1.84 &  2.35 &  2.65 &  2.90 &  3.15 &  3.43 &  4.08 \\
2000  &  1.06 &  1.20 &  1.53 &  1.91 &  2.13 &  2.32 &  2.48 &  2.68 &  3.08 \\
5000  &  0.95 &  1.07 &  1.30 &  1.56 &  1.69 &  1.81 &  1.91 &  2.04 &  2.30 \\
10000 &  0.96 &  1.05 &  1.23 &  1.41 &  1.50 &  1.58 &  1.65 &  1.75 &  1.92 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
```

```{r empirical-central-normality-critical-statistics, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
cap <- paste0(
  "Critical test statistic $\\tau_{\\alpha, n, \\kappa = 0.70}$ of the (empirical) ",
  "central normality test as function of $n$ for several values of significance level $\\alpha$. ",
  "The critical test statistics for central normality test are determined using fully normal data, ",
  "whereas the statistics for the empirical variant are determined using centrally normal data, ",
  "i.e. with fully normal data where 10% of elements are replaced by outliers. ",
  "emp: empirical"
)

.plot_test_statistic_tau(
  plot_theme = plot_theme,
  manuscript_dir = manuscript_dir
)

```

# Appendix F: Normalisation before transformation

An alternative to location- and scale-invariant transformations is normalising feature distributions prior to conventional transformations.
We assessed two normalisation methods were used:
1. z-standardisation, i.e. $x^{\prime}_{i} = \left(x_i - \mu \right) / \sigma$, with $\mu$ and $\sigma$ the mean and standard deviation of feature $\mathbf{X}$; 
and 2. robust scaling, i.e. $x^{\prime}_{i} = \left(x_i - \text{median}\left(\mathbf{X}\right) \right) / \text{IQR}\left(\mathbf{X}\right)$, with $\text{IQR}$ representing the interquartile range of feature $\mathbf{X}$.

Table \ref{tab:normalisation-before-transformation-appendix} shows residual errors after transformation to normality of the five features from real-world datasets presented previously in Appendix D and the main manuscript.
Table \ref{tab:normalisation-before-transformation-lambda-appendix} shows the corresponding values for transformation parameter $\lambda$, 
\ref{tab:normalisation-before-transformation-p-value-appendix} shows p-values of the empirical central normality test.
Normalisation prior to transformation prevents unexpectedly high $\lambda$ values in the presence of large location shifts, as can be observed for the latitude feature.
Normalisation prior to transformation, however, does not consistently improve residual errors after transformation. 
Interestingly, use of normalisation prior to transformation with robust power transformation using the method by Raymaekers and Rousseeuw [@Raymaekers2024-zf] led to higher residual errors for arterial wall thickness, fuel efficiency and latitude features.
This is similar to observations by @Raymaekers2024-zf, who observed that their method can emphasize values of outliers after transformation.
Location- and scale-invariant transformations have similar or lower residual errors compared to errors resulting from normalisation prior to transformation.
Not surprisingly, normalisation before transformation does not noticeably improve central normality, as assessed using the empirical central normality test. 

```{r include = FALSE, eval = FALSE}

data <- .assess_standardisation_before_transformation()
data <- data[, mget(c("residuals", "lambda", "p_value", "transformation", "dataset"))]
data[, "residuals" := round(residuals, digits = 1L)]
data[, "lambda" := round(lambda, digits = 1L)]
data[, "p_value" := round(p_value, digits = 3L)]
data.table::dcast(data, transformation ~ dataset, value.var = "residuals")
data.table::dcast(data, transformation ~ dataset, value.var = "lambda")
data.table::dcast(data, transformation ~ dataset, value.var = "p_value")
```

```{=latex}
\begin{table}
\begin{center}
\caption{Residual errors for features from real-world datasets after Yeo-Johnson transformation to normality. 
Several transformation methods include normalisation before tranformation, indicated by z-score normalisation (norm.) or robust scaling. 
AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
\label{tab:normalisation-before-transformation-appendix}
\begin{tabular}{l | r r r r r}

\toprule
feature & age & AWT & FE & latitude & PBM \\

\midrule
none & 16.5 & 110.1	& 54.5 & 328.4 & 48.0 \\
conventional & 11.5 & 30.0 & 55.3 & 319.0 & 32.2 \\
conventional (z-score norm.) & 11.5 & 19.3 & 49.0 & 326.2 & 33.3 \\
conventional (robust scaling) & 11.3 & 31.8 & 53.3 & 324.5 & 32.2 \\
Raymaekers-Rousseeuw & 11.5 & 136.7 & 47.7 & 315.1 & 32.2 \\
Raymaekers-Rousseeuw (z-score norm.) & 13.2 & 214.5 & 57.4 & 326.2 & 33.3 \\
Raymaekers-Rousseeuw (robust scaling) & 11.7 & 281.3 & 57.0 & 324.5 & 32.0 \\
invariant & 8.8 & 12.2 & 44.0 & 326.4 & 26.8 \\
robust invariant & 10.6 & 27.1 & 49.5 & 310.6 & 23.1 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
```

```{=latex}
\begin{table}
\begin{center}
\caption{Transformation parameter $\lambda$ for features from real-world datasets after Yeo-Johnson transformation to normality. 
Several transformation methods include normalisation before tranformation, indicated by z-score normalisation (norm.) or robust scaling. 
AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
\label{tab:normalisation-before-transformation-lambda-appendix}
\begin{tabular}{l | r r r r r}

\toprule
feature & age & AWT & FE & latitude & PBM \\

\midrule
conventional & 2.0 & -0.7 & -0.1 & 62.1 & -0.5 \\
conventional (z-score norm.) & 1.2 & -1.7 & -0.1 & 1.3 & 0.6 \\
conventional (robust scaling) & 1.3 & 0.0 & 0.2 & 1.4 & 0.4 \\
Raymaekers-Rousseeuw & 2.0 & 1.1 & 0.8 & 95.4 & -0.5 \\
Raymaekers-Rousseeuw (z-score norm.) & 1.3 & 1.4 & 1.0 & 1.3 & 0.6 \\
Raymaekers-Rousseeuw (robust scaling) & 1.4 & 1.3 & 1.0 & 1.4 & 0.4 \\
invariant & 1.3 & 0.2 & -1.3 & 1.5 & 0.5 \\
robust invariant & 1.4 & -0.6 & 0.9 & 1.4 & 0.4 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
```

```{=latex}
\begin{table}
\begin{center}
\caption{
p-values of empirical central normality tests for features from real-world datasets after Yeo-Johnson transformation to normality.
Several transformation methods include normalisation before tranformation, indicated by z-score normalisation (norm.) or robust scaling. 
AWT: arterial wall thickness; FE: fuel efficiency; PBM: penguin body mass}
\label{tab:normalisation-before-transformation-p-value-appendix}
\begin{tabular}{l | r r r r r}

\toprule
feature & age & AWT & FE & latitude & PBM \\

\midrule
none                                  & 0.712 & 0.864 & 0.577     & $< 0.001$ & $< 0.001$ \\
conventional                          & 0.936 & 0.017 & $< 0.001$ & $< 0.001$ & 0.058 \\
conventional (z-score norm.)          & 0.860 & 0.058 & $< 0.001$ & $< 0.001$ & 0.054 \\
conventional (robust scaling)         & 0.881 & 0.001 & $< 0.001$ & $< 0.001$ & 0.076 \\
Raymaekers-Rousseeuw                  & 0.936 & 0.905 & 0.320     & $< 0.001$ & 0.058 \\
Raymaekers-Rousseeuw (z-score norm.)  & 0.605 & 0.922 & 0.635     & $< 0.001$ & 0.054 \\
Raymaekers-Rousseeuw (robust scaling) & 0.830 & 0.948 & 0.637     & $< 0.001$ & 0.085 \\
invariant                             & 0.948 & 0.231 & $< 0.001$ & $< 0.001$ & 0.184 \\
robust invariant                      & 0.797 & 0.182 & 0.352     & $< 0.001$ & 0.397 \\
\bottomrule
\end{tabular}
\end{center}
\end{table}
```

# References
